\section{Motion Planning Using the Double Integrator Model} \label{sec:motion_planning_using_point_mass}

This section develops a motion planning framework using the double integrator model from \ref{subsec:point_mass_model}, transitioning from a global
coordinate representation to a Frenet frame formulation, as introduced in \ref{subsec:curve_following_coordinate_system}.
The Frenet frame provides a more structured approach for path tracking along a predefined road, accounting for curvature and alignment errors.
Our formulation builds on the work of Eilbrecht et al.
\cite{eilbrecht_challenges_2020} and systematically addresses motion dynamics, constraints, and control strategies.

\subsection{Overview of the Motion Planning Framework}

To effectively model and control vehicle motion, we employ a structured approach that systematically integrates kinematics, dynamics, constraints,
and control transformations.

We begin by defining the \textbf{coordinate transformation and kinematics}, introducing the curvature $C(s)$ of the reference path and the alignment
error $\xi = \psi - \theta$.
These definitions allow us to derive the first- and second-order kinematic equations, which describe how the vehicle's body-fixed velocity components
influence its evolution in Frenet coordinates.

Next, we formulate the \textbf{system dynamics} within the Frenet frame, defining the state vector and control inputs.
By incorporating the effects of curvature and alignment error, we capture both longitudinal and lateral dynamics within our model.

To address the nonlinearities introduced by the curvature terms, we employ \textbf{feedback linearization}.
This process involves assuming alignment to simplify the nonlinear dynamics, enabling the introduction of artificial control variables that linearize
the system.

To ensure that the planned trajectories remain within physical and operational limits, we carefully handle \textbf{constraint formulation}.
We derive bounds on velocity and acceleration and map them from the body-fixed frame to the Frenet frame, thereby enforcing vehicle and road
constraints within the planning model.

A key challenge in motion planning is the presence of \textbf{non-convexities} introduced by curvature-dependent constraints.
To address this, we use \textbf{quantifier elimination} techniques to obtain convex inner approximations of the feasible set.
We explore two approaches:
\begin{itemize}
	\item Interval Fitting, which provides a computationally efficient, box-shaped approximation of the constraint set.
	\item Cylindrical Algebraic Decomposition (CAD), a method from computer algebra that decomposes space into cylindrical cells to eliminate quantifiers while preserving logical equivalence.
\end{itemize}

Once an optimal trajectory is determined, we establish a \textbf{control transformation} that maps the optimized state and control variables to
physical vehicle inputs.
This step derives the necessary steering angle and longitudinal acceleration, ensuring compatibility with standard vehicle controllers.

Next, we provide the \textbf{exact discretization of the double integrator model}, ensuring an accurate transition from continuous to discrete-time
dynamics.
By leveraging matrix exponentials, this formulation preserves system behavior over discrete time steps.

Finally, we present the \textbf{complete motion planning model} as a formalized representation.
This encapsulates the system's state-space dynamics, control inputs, constraints, and transformation mappings.

\subsection{Coordinate Transformation and Kinematics}

In this subsection, we establish the foundation for our motion planning framework by describing the geometry of the reference path and deriving the
vehicle's kinematic equations in the Frenet frame.

First, let \(\theta(s)\) denote the tangent angle at an arc length \(s\) along the reference path.
The curvature, \(C(s)\), quantifies the rate of change of this tangent angle with respect to \(s\) and is defined as:
\begin{equation}
	C(s) := \frac{d\theta}{ds}.
\end{equation}

Next, consider the vehicle's orientation, \(\psi\).
To measure how much the vehicle deviates from following the road's direction, we define the alignment error \(\xi\) as:
\begin{equation}
	\xi := \psi - \theta.
\end{equation}
This error quantifies the difference between the vehicle's heading and the path's tangent direction.

Using these definitions and standard coordinate transformation techniques \cite{eilbrecht_challenges_2020}, we can derive the vehicle's motion
dynamics in the Frenet frame.
In this framework, the velocity components in the vehicle's body-fixed frame are directly related to the time derivatives of the Frenet coordinates.

\paragraph{First-Order Kinematics}\label{par:first_order_kinematics}
The following equations describe how the vehicle's position evolves along the path:
\begin{align}
	\dot{s}\,(1 - n\,C(s)) & = v_x\cos{\xi} - v_y\sin{\xi}, \label{eq:first_derivative_long} \\
	\dot{n}                & = v_x\sin{\xi} + v_y\cos{\xi}, \label{eq:first_derivative_lat}
\end{align}
where:
\begin{itemize}
	\item \(s\) is the longitudinal position along the reference path,
	\item \(n\) represents the lateral deviation from the path,
	\item \(v_x\) and \(v_y\) are the velocity components in the body-fixed frame.
\end{itemize}
Note that the term \(1 - n\,C(s)\) adjusts the longitudinal progress to account for the path's curvature.

\paragraph{Acceleration Dynamics}\label{par:acceleration_dynamics}
To capture the dynamics of acceleration in the Frenet frame, we introduce transformed acceleration components \(a_{x,tn}\) and \(a_{y,tn}\).
These dynamics are given by:
\begin{align}
	a_{x,tn} & = (a_x - v_y\,\dot{\psi})\cos{\xi} - (a_y + v_x\,\dot{\psi})\sin{\xi}, \label{eq:second_derivative_long} \\
	a_{y,tn} & = (a_x - v_y\,\dot{\psi})\sin{\xi} + (a_y + v_x\,\dot{\psi})\cos{\xi}, \label{eq:second_derivative_lat}
\end{align}
with the following definitions:
\begin{align}
	a_{x,tn} & := \ddot{s}\,(1 - n\,C(s)) - 2\dot{n}\,C(s)\dot{s} - n\,C'(s)\dot{s}^2, \label{def:axtn} \\
	a_{y,tn} & := \ddot{n} + C(s)\dot{s}^2\,(1 - n\,C(s)). \label{def:aytn}
\end{align}
These equations illustrate how the vehicle's acceleration in the Frenet frame is influenced by both its inherent dynamics (through \(\ddot{s}\) and
\(\ddot{n}\)) and the geometry of the reference path (through \(C(s)\) and its derivative \(C'(s)\)).

In summary, this subsection defines the key geometric parameters and derives the kinematic equations necessary for representing vehicle motion in the
Frenet frame.
These results set the stage for the subsequent development of the full motion planning model.

\subsection{System Dynamics Formulation}

We can now formalize the motion planning model in the Frenet frame by defining the state and control input vectors.
The state vector, denoted by \(x_{di}\), captures the vehicle's position, alignment error, and their corresponding velocities:
\begin{equation}
	x_{di} = \begin{bmatrix}
		s       \\
		n       \\
		\xi     \\
		\dot{s} \\
		\dot{n} \\
		\dot{\psi}
	\end{bmatrix},
\end{equation}
where:
\begin{itemize}
	\item \(s\) is the longitudinal position along the reference path,
	\item \(n\) is the lateral deviation,
	\item \(\xi\) is the alignment error (\(\psi - \theta\)),
	\item \(\dot{s}\), \(\dot{n}\), and \(\dot{\psi}\) are the corresponding time derivatives.
\end{itemize}

The control inputs are represented by the vector \(u_{di}\), which comprises the body-fixed accelerations:
\begin{equation}
	u_{di} = \begin{bmatrix}
		a_x \\
		a_y \\
		a_\psi
	\end{bmatrix}.
\end{equation}

Using the kinematic relationships derived earlier in \nameref{par:first_order_kinematics} and \nameref{par:acceleration_dynamics}, the dynamics of the double integrator model in the Frenet frame can be expressed as:
\begin{equation}
	\label{eq:frenet_frame_pm_dynamics_0}
	f_{di}(x_{di}, u_{di}) = \begin{bmatrix}
		\dot{s}                                                                                \\
		\dot{n}                                                                                \\
		\dot{\psi} - C(s)\dot{s}                                                               \\
		\displaystyle \frac{a_{x,tn} + 2\dot{n}\,C(s)\dot{s} + n\,C'(s)\dot{s}^2}{1 - n\,C(s)} \\
		a_{y,tn} - C(s)\dot{s}^2(1 - n\,C(s))                                                  \\
		a_\psi
	\end{bmatrix}.
\end{equation}

This formulation captures the vehicle's motion along a curved path by incorporating both its intrinsic dynamics and the influence of road curvature.
It is important to note that the curvature terms \(C(s)\) and \(C'(s)\) introduce non-convexities into the system dynamics.
We address these challenges in subsequent sections through feedback linearization.

\subsection{Linearization via Feedback Control} \label{subsec:constraints}
To simplify the nonlinear dynamics introduced by the curvature terms, we first decouple the body-fixed inputs in
\eqref{eq:frenet_frame_pm_dynamics_0}.
By substituting \eqref{eq:frenet_frame_pm_dynamics_0} the accelerations in the Frenet frame with their body-fixed relations from
\eqref{eq:second_derivative_long} and \eqref{eq:second_derivative_lat} in \eqref{eq:frenet_frame_pm_dynamics_0}, we obtain:
\begin{align}
	\frac{(a_x - v_y\,\dot{\psi})\cos{\xi} - (a_y + v_x\,\dot{\psi})\sin{\xi} + 2\dot{n}\,C(s)\dot{s} + n\,C'(s)\dot{s}^2}{1 - n\,C(s)} \\
	(a_x - v_y\,\dot{\psi})\sin{\xi} + (a_y + v_x\,\dot{\psi})\cos{\xi} - C(s)\dot{s}^2(1 - n\,C(s))
\end{align}
We observe that both entries contain the body-fixed control inputs \(a_x\) and \(a_y\).
To simplify the model, we make the following key assumption.

\subsubsection{Assumption: Alignment Error} \label{subsubsec:alignment_error}
We assume that the vehicle is always aligned with the road, i.e.,
\begin{equation}
	\xi = \psi - \theta = 0.
\end{equation}
This assumption leads to several useful simplifications:
\begin{itemize}
	\item The body-fixed accelerations are directly given by the transformed accelerations:
	      \[
		      [a_x,\, a_y] = [a_{x,tn},\, a_{y,tn}].
	      \]
	\item The rate of change of the vehicle's orientation becomes:
	      \[
		      \dot{\psi} = \dot{\theta} = C(s)\,\dot{s},
	      \]
	      where \(C(s)=\frac{d\theta}{ds}\).
	\item The yaw acceleration satisfies:
	      \[
		      a_\psi = \ddot{\psi} = \ddot{\theta} = C'(s)\,\dot{s}^2 + C(s)\,\ddot{s}.
	      \]
\end{itemize}
Although this assumption fixes the orientation to the road, the vehicle is still permitted lateral movement via lateral acceleration.

With this alignment assumption, the dynamics become affine in \(a_{x,tn}\) and \(a_{y,tn}\).
This property allows us to introduce artificial control inputs that will fully linearize the system.

\subsubsection{Introducing Artificial Control Inputs}

Define the artificial control input vector \(\tilde{u}_{di}\) as:
\begin{equation}
	\label{eq:artificial_controls}
	\tilde{u}_{di} := \begin{bmatrix}
		u_t \\
		u_n
	\end{bmatrix} :=
	\begin{bmatrix}
		\displaystyle \frac{a_{x,tn} + 2\dot{n}\,C(s)\dot{s} + n\,C'(s)\dot{s}^2}{1 - n\,C(s)} \\
		a_{y,tn} - C(s)\dot{s}^2(1 - n\,C(s))
	\end{bmatrix}.
\end{equation}
These artificial control inputs will be used to linearize the system dynamics, a process known as feedback linearization.
This technique is thoroughly explained in \cite{khalil_nonlinear_2002}.
In the following section, we will briefly introduce the concept of feedback linearization before applying it in our motion planning framework.

\subsubsection{Supplementary Background: Feedback Linearization}

Feedback linearization is a nonlinear control technique that transforms a nonlinear system into an equivalent linear system by means of a suitable
change of variables and state-feedback law.
Consider a general nonlinear system of the form:
\begin{equation}
	\dot{x} = f(x) + G(x)\,u,
\end{equation}
where
\begin{itemize}
	\item \(x \in \mathbb{R}^n\) is the state vector,
	\item \(f(x)\) represents the system dynamics,
	\item \(G(x)\) is the input matrix,
	\item \(u \in \mathbb{R}^m\) is the control input.
\end{itemize}
For feedback linearization, it is typical to assume that the system is fully actuated, i.e., \[ \text{rank}\bigl(G(x)\bigr) = n, \] ensuring that
\(G(x)\) is invertible in the region of interest.
Under this assumption, one can define a new control input \(v\) such that:
\begin{equation}
	u = G(x)^{-1}\,\bigl[v - f(x)\bigr].
\end{equation}
By canceling the nonlinear dynamics \(f(x)\), the new input \(v\) governs an equivalent linear system that can be addressed with standard linear
control techniques.

\subsubsection{Resulting Simplified Model}\label{subsubsec:resulting_simplified_model}

Returning to our model, we can use the artificial control inputs \(\tilde{u}_{di}\) to linearize the system dynamics.
Given our assumption of no alignment error \ref{subsubsec:alignment_error}, the state variables can be simplified by removing the orientation
\(\psi\), as it is fixed to the road.
The reduced state vector is:
\[
	\tilde{x}_{di} = \begin{bmatrix} s, & n, & \dot{s}, & \dot{n} \end{bmatrix}^T,
\]
and the new artificial control inputs are:
\[
	\tilde{u}_{di} = \begin{bmatrix} u_t, & u_n \end{bmatrix}^T.
\]
Under these definitions, the system dynamics simplify to:
\begin{equation}
	\label{eq:pm_final_dynamics}
	\tilde{f}_{di}(\tilde{x}_{di}, \tilde{u}_{di}) = \begin{bmatrix}
		\dot{s} \\
		\dot{n} \\
		u_t     \\
		u_n
	\end{bmatrix}.
\end{equation}

With the dynamics now expressed in a simplified, linear form, our next task is to ensure that the planned trajectories adhere to both the vehicle's
physical limitations and the geometric constraints of the road.
In the following section, we establish the coupling constraints between the state variables and control inputs for our discrete-time optimal
trajectory planning problem (see \eqref{eq:coupling_constraints}) based on \cite{eilbrecht_challenges_2020}.

\subsection{Constraint Handling}

This subsection addresses the constraints imposed by the vehicle's physical limits and the road geometry.
We first define the constraints on state variables and control inputs in the body-fixed frame and then map them to the Frenet frame.
This mapping allows us to formulate the overall coupling constraint set that must be satisfied during trajectory planning.
We will also discuss the challenge of non-convexity arising from these coupling constraints and formulate the problem of finding a convex
under-approximation.

Let \(\square\) denote any planning variable.
For each variable, we define constant upper and lower bounds during planning, denoted by \(\overline{\square}\) and \(\underline{\square}\),
respectively.
For example, in the vehicle's body-fixed frame the velocity constraints are expressed as:
\begin{align}
	\underline{v_x} \leq v_x \leq \overline{v_x}, \\
	\underline{v_y} \leq v_y \leq \overline{v_y}.
\end{align}

By applying the first-order kinematics from \eqref{eq:first_derivative_long} and \eqref{eq:first_derivative_lat} with the alignment error set to \(\xi=0\), these velocity bounds translate into the Frenet frame as follows:
\begin{align}
	\underline{v_x} \leq \dot{s}\,(1 - n\,C(s)) \leq \overline{v_x}, \\
	\underline{v_y} \leq \dot{n} \leq \overline{v_y}.
\end{align}
Here, \(\dot{s}\) represents the longitudinal speed adjusted by the term \((1 - n\,C(s))\) to account for the curvature of the road, and \(\dot{n}\)
is the lateral speed.

For acceleration constraints, we typically define two types:
\begin{itemize}
	\item A norm constraint that limits the overall acceleration, ensuring that the combined longitudinal and lateral accelerations lie within a circle of radius \(c\), similar to \eqref{eq:acceleration_constraint_preliminaries_di}:
	      \begin{equation}
		      a_x^2 + a_y^2 \leq c.
	      \end{equation}
	\item Individual bounds on the longitudinal and lateral accelerations:
	      \begin{align}
		      \underline{a_x} \leq a_x \leq \overline{a_x}, \\
		      \underline{a_y} \leq a_y \leq \overline{a_y}.
	      \end{align}
\end{itemize}

To map these acceleration constraints to the Frenet frame, we use the definition \eqref{eq:artificial_controls} of our artificial variables
\eqref{eq:artificial_controls} and the first implication $a_x=a_{x,tn}$ and $a_y=a_{y,tn}$ from our alignment error assumption
\ref{subsubsec:alignment_error}.
These equations allow us to establish a mapping that relates the state variables and artificial control inputs \((\tilde{x}_{di}, \tilde{u}_{di})\)
to the body-fixed accelerations by solving for $a_x$ and $a_y$.
This mapping is defined as:
\begin{equation}
	\label{def:g}
	g(\tilde{x}_{di}, \tilde{u}_{di}) :=
	\begin{bmatrix}
		(1 - n\,C(s))\,u_t - \bigl(2\dot{n}\,C(s)\dot{s} + n\,C'(s)\dot{s}^2\bigr) \\
		u_n + C(s)\,\dot{s}^2\,(1 - n\,C(s))
	\end{bmatrix}
	=
	\begin{bmatrix}
		a_x \\
		a_y
	\end{bmatrix}.
\end{equation}
Substituting this mapping into the individual acceleration constraints, we derive the following bounds in the Frenet frame:
\begin{align}
	\begin{bmatrix}
		\underline{a_x} \\[2mm] \underline{a_y}
	\end{bmatrix} \leq g(\tilde{x}_{di}, \tilde{u}_{di}) \leq \begin{bmatrix}
		                                                          \overline{a_x} \\[2mm] \overline{a_y}
	                                                          \end{bmatrix} \\[2mm]
	\|g(\tilde{x}_{di}, \tilde{u}_{di})\|^2 \leq c.
\end{align}

Next, we need to impose limits on the yaw rate and yaw acceleration.
The yaw rate is defined as \(C(s)\,\dot{s}\), and the yaw acceleration is given by \(C'(s)\,\dot{s}^2 + C(s)\,u_t\), as derived from the second and
third implications of our alignment error assumption \ref{subsubsec:alignment_error}.
\begin{align}
	\underline{\dot{\psi}} \leq C(s)\,\dot{s} \leq \overline{\dot{\psi}}, \\
	\underline{a_{\psi}} \leq C'(s)\,\dot{s}^2 + C(s)\,u_t \leq \overline{a_{\psi}}.
\end{align}

Thus, we have successfully modeled the physical limits of the vehicle in the Frenet frame.
The constraints arising from the road topology can be represented using the Frenet frame coordinates from our state variables, \(s\) and \(n\).
To adhere to DCP rules, we allow the lateral range to depend on the arc length \(s\).
This is achieved by defining the bounds \(\underline{n}(s)\) and \(\overline{n}(s)\) for the lateral deviation, where \(\underline{n}(s)\) is convex
in \(s\) and \(\overline{n}(s)\) is concave in \(s\).
The arc length \(s\) is constrained by the constant bounds \(\underline{s}\) and \(\overline{s}\).

Combining all introduced constants, the overall coupling constraint set \(\mathcal{C}\) is defined as:
\begin{equation}
	\mathcal{C} := \left\{
	\begin{bmatrix} \tilde{x}_{di} \\ \tilde{u}_{di} \end{bmatrix} \; \middle|\;
	\begin{aligned}
		 & \underline{s} \leq s \leq \overline{s},                                                        \\
		 & \underline{n}(s) \leq n \leq \overline{n}(s),                                                  \\
		 & \underline{v_x} \leq \dot{s}\,(1 - n\,C(s)) \leq \overline{v_x},                               \\
		 & \underline{v_y} \leq \dot{n} \leq \overline{v_y},                                              \\
		 & \underline{\dot{\psi}} \leq C(s)\,\dot{s} \leq \overline{\dot{\psi}},                          \\
		 & \underline{a_{\psi}} \leq C'(s)\,\dot{s}^2 + C(s)\,u_t \leq \overline{a_{\psi}},               \\
		 & \begin{bmatrix}
			   \underline{a_x} \\[2mm] \underline{a_y}
		   \end{bmatrix} \leq g(\tilde{x}_{di}, \tilde{u}_{di}) \leq \begin{bmatrix}
			                                                             \overline{a_x} \\[2mm] \overline{a_y}
		                                                             \end{bmatrix}, \\
		 & \|g(\tilde{x}_{di}, \tilde{u}_{di})\|^2 \leq c.
	\end{aligned}
	\right\}.
\end{equation}

The constraint set \(\mathcal{C}\) is highly non-convex, primarily due to the curvature terms \(C(s)\) and \(C'(s)\) and their nonlinear interaction
with the state and control inputs.
To handle this non-convexity, we seek to derive a convex inner approximation of the feasible set \(\mathcal{C}\).
This objective is defined here and addressed in the following section.

\subsubsection{Problem Definition: Finding an Inscribed Polytope}
\label{problem:inscribed_polytope}

To handle the non-convexity, our approach is to approximate \(\mathcal{C}\) with an inscribed polytope \(\hat{C}\) that is convex.
Formally, we seek to determine:
\begin{equation}
	\hat{C} = \left\{ \begin{bmatrix}
		\tilde{x}_{di} \\[2mm] \tilde{u}_{di}
	\end{bmatrix} \; \middle|\;
	N \begin{bmatrix}
		\tilde{x}_{di} \\[2mm] \tilde{u}_{di}
	\end{bmatrix} \leq b
	\right\} \subseteq \mathcal{C},
\end{equation}
where \(N\) and \(b\) represent a set of linear inequalities whose intersection forms the polytope.
In the following section, we will demonstrate how to address this problem.

\subsection{Convex Approximation via Quantifier Elimination}
\input{content/methodology/convex_apporximation_using_qe.tex}

\subsection{Determining the Steering Angle} \label{subsec:determining_the_steering_angle}

Typically, a vehicle is controlled through throttle, brakes, and a steering angle.
To incorporate these controls, we need to move away from visualizing our model as a box aligned with the road.
Instead, we will treat the model as a point and define its orientation based on its velocity.
We set $v_y = 0$ and $a_y=0$ to reflect that lateral motion arises solely from steering, thereby keeping the lateral velocity state zero and
capturing lateral movement through changes in heading rather than a separate lateral speed.
Using the equations \eqref{eq:first_derivative_long} and \eqref{eq:first_derivative_lat} with $v_y = 0$, we can solve for $v_x$.
\begin{equation}
	v := v_x = \sqrt{(1-nC(s))^2\dot{s}^2 + \dot{n}^2}
\end{equation}
Dividing $\dot{n}$ by $\dot{s}$ combined with  \eqref{eq:first_derivative_long} and \eqref{eq:first_derivative_lat} yields:
\begin{equation}
	\frac{\dot{n}}{\dot{s}} = (1-nC(s))\tan(\xi) = (1-nC(s))\tan(\psi - \theta)
\end{equation}
which we can solve for $\psi$ to get the orientation of the vehicle.
\begin{equation}
	\psi = \theta + \arctan\left(\frac{\dot{n}}{\dot{s}(1-nC(s))}\right)
\end{equation}

Using the state variables and $g$ from \eqref{def:g}, we can calculate $a_{x,tn}$ and $a_{y,tn}$ from \eqref{def:axtn} and \eqref{def:aytn},
respectively.
By substituting these values into equations \eqref{eq:second_derivative_long} and \eqref{eq:second_derivative_lat}, and setting $a_y = 0$, we can
determine the longitudinal acceleration and the change in orientation.
Additionally, we assume $|\xi| \leq \frac{\pi}{2}$ to ensure that $\cos{\xi} \neq 0$.
\begin{align}
	\dot{\psi} = \frac{a_{y,tn} - \tan(\xi) a_{x,tn}}{v (\tan(\xi) \sin(\xi) + \cos(\xi))} \\
	a := a_x = \frac{a_{x,tn} + v \dot{\psi} \sin(\xi)}{\cos{\xi}}
\end{align}
Our bicycle models \eqref{eq:dpsi_steering_angle} enables us to calculate the steering angle.
\begin{equation}
	\delta = \arctan(l_{wb}\frac{\dot{\psi}}{v})
\end{equation}
With those equations, we can define a transformation.
\begin{equation}
	T(\tilde{x}_{di}, \tilde{u}_{di}) = [p_x, p_y, \psi, \dot{\psi}, v, a, \delta] \label{eq:pm_state_transformation} \end{equation}

\subsection{Exact Discretization of the Double Integrator Model}

Using the simplified state and control representation from
Section~\ref{subsubsec:resulting_simplified_model}, the system dynamics \eqref{eq:pm_final_dynamics} are discretized using the matrix exponential
method \cite{kailath_linear_1980, ogata_modern_2010}.
The discrete-time system is formulated as:

\begin{equation}
	\label{eq:discrete_time_dynamics_di}
	\tilde{x}_{di, k+1} = A_d \tilde{x}_{di, k} + B_d \tilde{u}_{di, k} =: f_{d, di}(\tilde{x}_{di, k}, \tilde{u}_{di, k}, \Delta t),
\end{equation}

where the exact discretization is given by:

\begin{equation}
	A_d = e^{A \Delta t}, \quad B_d = \left( \int_0^{\Delta t} e^{A \tau} d\tau \right) B.
\end{equation}

Using the closed-form solution for the matrix exponential \cite{noauthor_matrix_nodate}, we obtain:

\begin{equation}
	A_d = \begin{bmatrix}
		1 & 0 & \Delta t & 0        \\
		0 & 1 & 0        & \Delta t \\
		0 & 0 & 1        & 0        \\
		0 & 0 & 0        & 1
	\end{bmatrix},
	\quad
	B_d = \begin{bmatrix}
		\frac{\Delta t^2}{2} & 0                    \\
		0                    & \frac{\Delta t^2}{2} \\
		\Delta t             & 0                    \\
		0                    & \Delta t
	\end{bmatrix}.
\end{equation}

This formulation ensures that the discrete-time representation accurately follows the continuous-time system over a sampling interval \( \Delta t \),
preserving dynamic consistency while enabling efficient convex optimization.
The resulting discrete-time model will be used in the subsequent trajectory planning formulation.

\subsection{Final Model Representation} \label{subsec:pm_resulting_model}

Our final model is represented by the following tuple.
\begin{equation}
	M_{pm} = (\tilde{x}_{di}, \tilde{u}_{di}, f_{d, di}, \hat{C}, T)
	\label{model:point_mass}
\end{equation}
