\chapter{Introduction} \label{ch:intro}

The \textit{student starter kit} serves as guidance for students writing a thesis. This Latex document contains some advice on writing and can be used as template for your thesis. 
In addition, this starter kit contains a template for presentations, standardized grade sheets used to transparently mark your work, and a template for submitting your files (see Sec.~\ref{sec:submission}). As a short summary, the kit also contains two text files: a checklist for your thesis and one for your presentation; please adhere to them!

The student starter kit is only to be used within the Professorship Cyber-Physical Systems. The kit has been created by Markus Koschi\footnote{\href{mailto:markus.koschi@tum.de}{\texttt{markus.koschi@tum.de}}}. This latex template is based on the template by Florian Walch and contributors\footnote{\url{https://github.com/fwalch/tum-thesis-latex}}. The advice on format and style are from Carmella Schürmann. The advice on the literature review are from Matthias Althoff. 

Further reading is referenced where applicable, and general information provided by TUM is available in German\footnote{\url{https://www.tum.de/studium/studienabschluss/abschlussarbeit/tipps-und-tricks/}} and English\footnote{\url{https://www.tum.de/nc/en/studies/completing-your-studies/theses/tips-and-tricks/}}. The advice given by Elmar Juergens, a former PhD student at TUM in software engineering, might be also useful for you: \url{https://thesisguide.org/}.


\section{Submission} \label{sec:submission}
For internal archiving, you are required to submit all source files of your thesis, presentation, and software in the given format (see the enclosed submission folder):
The parent folder should be named firstNameLastName and contain the following sub-folders:
\begin{itemize}
\item papers\_FirstNameLastName (folder containing studied publications)
\item presentation\_FirstNameLastName (folder for the pdf and source files of the presentation)
\item report\_FirstNameLastName (folder for the pdf and source files of the report)
\item software\_FirstNameLastName (folder for the source files of the created software)
\end{itemize}
Please submit this folder at latest on the day of your presentation either via mail or personally by flash drive. The pdf of your thesis is required on the official day of submission. Note that you should aim at reproducible results, i.e., that your advisor can run your code and reproduce your results. Thus, providing a read-me and saving the simulation parameters in a script might be a good idea.

\section{Overview}
This document explains how to structure a thesis in Ch.~\ref{ch:format} and how to write well in Ch.~\ref{ch:Style}. Ch.~\ref{ch:latex} provides advice on how to use Latex, and Ch.~\ref{ch:lit_review} describes best practices to conduct a literature review.







